Lyman-$\alpha$ photons through rotating outflows

### 2018-03-31 paper/paper.tex
references

\begin{keywords}
galaxies: dwarf --- radiative transfer --- Methods: numerical
\end{keywords}

@@

Recent advances in instrumentation have revolutionized the kind of
kinematic studies that can be performed on galaxies with small angular
sizes on the sky.
For instance, spatially resolved spectra on compact dwarf galaxies
have measured clear signs of gas showing pure rotation kinematics
\citep{2015A&A...577A..21C,2017A&A...600A.125C}; recent mapping of
high redshift circumgalactic regions have also revealed kinematic
evidence for large scale rotation \citep{2018MNRAS.473.3907A}.
These systems with star formation, some neutral gas and low dust
contents naturally show a \lya emission line \citep{PartridgePeebles},
motivating the observational work to phenomenologically link
kinematics traced by H$\alpha$ to \lya spectra \citep[e.g.][]{Herenz2016}.

@@

However, the gas dynamics in Lyman Alpha Emitter (LAE) galaxies is
more complex than pure rotation.
The case for outflows as traced in the \lya\ line is much better
stablished.
In many observations the \lya line profiles has a single peak
redwards from the line's center, in other cases there is a double peak
but the peak on the red side is stronger
\citep[e.g.][]{2010ApJ...717..289S,Erb14,Trainor16}.   
These features been readily explained as the consequence of multiple
\lya photon scatterings through an homogeneous outflowing shell of
neutral Hydrogen
\citep{2006A&A...460..397V,Orsi12,2012ApJ...751...29Y,2015ApJ...812..123G}.

An study of the combined effects of outflows and rotation has not been
presented in the literature.
Here we present for the first time such as study. 
We study a simplified geometrical configuration corresponding to a
spherical gas cloud with symmetrical radial outflows and solid body
rotation. 
We base our modeling on a Monte-Carlo radiative transfer code.
Besides modeling the impact of joint rotation and outflows, we also
check to what extent the analytical model presented by
\cite{Garavito14} to explain the effects of rotation can also be
applied in our case.

>>>>

\begin{keywords}
radiative transfer --- Methods: numerical
\end{keywords}

@@

Recent advances in instrumentation have revolutionized the kind of
kinematic studies that can be performed on extra-galactic systems with small angular
sizes on the sky, revealing the presence of gas rotation on vastly
different physical scales.
For instance, spatially resolved spectra on compact dwarf galaxies
have measured clear signs of gas showing pure rotation kinematics
\citep{2015A&A...577A..21C,2017A&A...600A.125C}; recent mapping of
high redshift circumgalactic regions have also revealed kinematic
evidence for large scale rotation \citep{2018MNRAS.473.3907A}.
These systems with star formation, some neutral gas and low dust
contents naturally show a \lya emission line \citep{PartridgePeebles},
motivating the observational work to phenomenologically link
H$\alpha$ kinematics to \lya spectra \citep[e.g.][]{Herenz2016}.

@@

However, the gas dynamics in Lyman Alpha Emitter (LAE) galaxies is
more complex than pure rotation.
The case for outflows as traced in the \lya line is much better
stablished.
In many observations the \lya line profiles has a single peak
redwards from the line's center, in other cases there is a double peak
but the peak on the red side is stronger
\citep[e.g.][]{2010ApJ...717..289S,Erb14,Trainor16}.   
These features been readily explained as the consequence of multiple
\lya photon scatterings through an homogeneous outflowing shell of
neutral Hydrogen
\citep{2006A&A...460..397V,Orsi12,2012ApJ...751...29Y,2015ApJ...812..123G}.

An study of the combined effects of outflows and rotation has not been
presented in the literature.
Here we present such a study. 
We investigate a simplified geometrical configuration corresponding to a
spherical gas cloud with symmetrical radial outflows and solid body
rotation. 
We base our modeling on a Monte-Carlo radiative transfer code.
Besides modeling the impact of joint rotation and outflows, we also
check to what extent the analytical model presented by
\cite{Garavito14} to explain the effects of rotation can also be
applied in our case.
